---
title: "Dissertation Chapter 4 Statistical Analysis Models"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
rm(list = ls())
```

## Chapter 4 BRAIN SYSTEM SEGREGATION EXPLAINS DISCREPANCIES BETWEEN AD PATHOLOGY AND DEMENTIA STATUS


```{r define covariates}
# define covariates for cross-sectional analyses:
# age, self-reported gender,education status, head motion (post-scrubbing FD)
cross_cov = 'Age+Gender+Educ_Group+FD_postscrub'

# define covariates for longitudinal analyses: 
# age at baseline, self-reported gender,education status, head motion (post-scrubbing FD) at baseline
long_cov = 'Age_bl+Gender+Educ_Group+FD_postscrub_bl'
```

All reported cross-sectional analyses controlled for the above cross-sectional analyses covariates. 
All reported longitudinal analyses controlled for the above longitudinal analyses covariates.

### 4.3.2 Cognitive healthy individuals who harbor amyloid-β have higher brain system segregation

1) Examine difference in brain system segregation between cognitively healthy and cognitively impaired individuals designated as Aβ+: 

```{r model 4.3.2.1}
# DV: whole-brain system segregation
# IV: CDR group (CDR = 0 vs. CDR > 0)

mod4.3.2.1 <- paste('wb_seg ~ CDR_Group',cross_cov,sep = "+") 

# Perform ANCOVA
car::Anova(lm(formula = mod4.3.2.1,data = data,contrasts=
              list(CDR_Group="contr.treatment",Gender = "contr.sum",Educ_Factor=c(-1,1))),type=3) # perform ANCOVA
```

- Examine difference in brain system segregation between cognitively healthy and cognitively impaired individuals designated as Aβ+, while controlling for Centiloids: 

```{r model 4.3.2.2}
# DV: whole-brain system segregation
# IV: CDR scores
# controlling for centiloids and the primary covariates

mod4.3.2.2 <- paste('wb_seg ~ CDR_Group+Centiloids',cross_cov,sep = "+") 

# Perform ANCOVA
car::Anova(lm(formula = mod4.3.2.2,data = data,contrasts=
              list(CDR_Group="contr.treatment",Gender = "contr.sum",Educ_Factor=c(-1,1))),type=3) # perform ANCOVA

```

2) Examine difference in brain system segregation between cognitively healthy and cognitively impaired individuals designated as Tau+: 

```{r model 4.3.2.3}
# DV: whole-brain system segregation
# IV: CDR group (CDR = 0 vs. CDR > 0)

mod4.3.2.3 <- paste('wb_seg ~ CDR_Group',cross_cov,sep = "+") 

# Perform ANCOVA
car::Anova(lm(formula = mod4.3.2.3,data = data,contrasts=
              list(CDR_Group="contr.treatment",Gender = "contr.sum",Educ_Factor=c(-1,1))),type=3) # perform ANCOVA

```

- Examine difference in brain system segregation between cognitively healthy and cognitively impaired individuals designated as tau+, while controlling for meta-temporal SUVR: 

```{r model 4.3.2.4}
# DV: whole-brain system segregation
# IV: CDR scores
# controlling for meta-temporal SUVR and the primary covariates

mod4.3.2.4 <- paste('wb_seg ~ CDR_Group+meta_temporal_suvr',cross_cov,sep = "+") 

# Perform ANCOVA
car::Anova(lm(formula = mod4.3.2.4,data = data,contrasts=
              list(CDR_Group="contr.treatment",Gender = "contr.sum",Educ_Factor=c(-1,1))),type=3) # perform ANCOVA

```


3) Examine difference in brain system segregation between cognitively healthy and cognitively impaired individuals designated as amyloid+ and tau+: 

```{r model 4.3.2.5}
# DV: whole-brain system segregation
# IV: CDR group (CDR = 0 vs. CDR > 0)

mod4.3.2.5 <- paste('wb_seg ~ CDR_Group',cross_cov,sep = "+") 

# Perform ANCOVA
car::Anova(lm(formula = mod4.3.2.5,data = data,contrasts=
              list(CDR_Group="contr.treatment",Gender = "contr.sum",Educ_Factor=c(-1,1))),type=3) # perform ANCOVA

```


### 4.3.3 The segregation of specific brain systems moderates the relationship between tau and cognitive dysfunction

1) Examine whether brain system segregation moderates the relationship between meta-temporal tau burden and CDR-SOB scores:

```{r model 4.3.3.1}

# DV: CDR-SOB
# IV: interaction between whole-brain system segregation and meta-temporal tau uptake
mod4.3.3.1 = paste('CDR_SOB ~ wb_seg*meta_temporal_suvr',cross_cov,sep = "+") 

#perform multiple linear regression
lm(formula = mod4.3.3.1,data = data,contrasts = list(Gender = "contr.sum",Educ_Group=c(-1,1))) 
```

2) Examine whether sensory-motor system segregation moderates the relationship between meta-temporal tau burden and CDR-SOB scores:

```{r model 4.3.3.2}
# DV: CDR-SOB
# IV: interaction between sensory-motor system segregation and meta-temporal tau uptake
mod4.3.3.2 = paste('CDR_SOB ~ sensory_seg*meta_temporal_suvr',cross_cov,sep = "+") 

#perform multiple linear regression
lm(formula = mod4.3.3.2,data = data,contrasts = list(Gender = "contr.sum",Educ_Group=c(-1,1))) 

```

3) Examine whether association system segregation moderates the relationship between meta-temporal tau burden and CDR-SOB scores:

```{r model 4.3.3.3}
# DV: CDR-SOB
# IV: interaction between association system segregation and meta-temporal tau uptake
mod4.3.3.3 = paste('CDR_SOB ~ asso_seg*meta_temporal_suvr',cross_cov,sep = "+") 

#perform multiple linear regression
lm(formula = mod4.3.3.3,data = data,contrasts = list(Gender = "contr.sum",Educ_Group=c(-1,1))) 

```

### 4.3.4 Brain system segregation relates to dementia severity, beyond the variance explained by AD-related pathology and measures of structural neurodegeneration

1) Examine the main effects of brain system segregation, cortical Aβ (i.e., Centiloids), meta-temporal tau deposition, mean grey matter cortical thickness and mean hippocampal volume on CDR scores: 

```{r model 4.3.4.1}
# DV: CDR scores
# IV: whole-brain system segregation, cortical Aβ (i.e., centiloids), meta-temporal tau deposition, mean grey matter cortical thickness and mean hippocampal volume 
mod4.3.4.1 = paste('CDR ~ Centiloids+meta_temporal_suvr+mean_cortical_thickness+hippo_volume+wb_seg',cross_cov,sep = "+") 

#perform multiple linear regression
lm(formula = mod4.3.4.1,data = data,contrasts = list(Gender = "contr.sum",Educ_Group=c(-1,1))) 

```

2) Examine the main effects of brain system segregation, cortical Aβ (i.e., Centiloids), meta-temporal tau deposition, mean grey matter cortical thickness and mean hippocampal volume on CDR-SOB scores: 

```{r model 4.3.4.2}
# DV: CDR-SOB scores
# IV: whole-brain system segregation, cortical Aβ (i.e., centiloids), meta-temporal tau deposition, mean grey matter cortical thickness and mean hippocampal volume 

mod4.3.4.2 = paste('CDR_SOB ~ Centiloids+meta_temporal_suvr+mean_cortical_thickness+hippo_volume+wb_seg',cross_cov,sep = "+") 

#perform multiple linear regression
lm(formula = mod4.3.4.2,data = data,contrasts = list(Gender = "contr.sum",Educ_Group=c(-1,1))) 

```

3) Examine the main effects of brain system segregation, cortical Aβ (i.e., Centiloids), meta-temporal tau deposition, mean grey matter cortical thickness and mean hippocampal volume on ADAS-cog scores: 

```{r model 4.3.4.3}
# DV: ADAS-cog scores
# IV: whole-brain system segregation, cortical Aβ (i.e., centiloids), meta-temporal tau deposition, mean grey matter cortical thickness and mean hippocampal volume 

mod4.3.4.3 = paste('ADAS_Cog ~ Centiloids+meta_temporal_suvr+mean_cortical_thickness+hippo_volume+wb_seg',cross_cov,sep = "+") 

#perform multiple linear regression
lm(formula = mod4.3.4.3,data = data,contrasts = list(Gender = "contr.sum",Educ_Group=c(-1,1))) 

```

### 4.3.5 Baseline brain system segregation predicts cognitive decline over time, beyond other AD brain measures

1) Examine the association between brain system segregation at baseline and change in cognitive status over time:

- Examine whether brain system segregation at baseline predicts change in CDR-SOB scores: 

```{r model 4.3.5.1}
# DV: CDR-SOB scores
# IV: whole brain system segregation at baseline
# Controlling for cortical Aβ (i.e., centiloids), meta-temporal tau deposition, mean grey matter cortical thickness and mean hippocampal volume at baseline, together with the longitudinal covariates

mod4.3.5.1 = paste('CDR_SOB ~ wb_seg_bl*time+Centiloids_bl+meta_temporal_suvr_bl+mean_cortical_thickness_bl+hippo_volume_bl',long_cov,"(1|time|Subject_ID)",sep = "+")

# perform linear mixed effect regression
lmer(formula = mod3.3.5.5,data = data,contrasts=list(Gender = "contr.sum",Educ_Group=c(-1,1)))
```

- Examine whether brain system segregation at baseline predicts change in CDR score: 

```{r model 4.3.5.2}
# DV: CDR scores
# IV: whole brain system segregation at baseline
# Controlling for cortical Aβ (i.e., centiloids), meta-temporal tau deposition, mean grey matter cortical thickness and mean hippocampal volume at baseline, together with the longitudinal covariates

mod4.3.5.2 = paste('CDR ~ wb_seg_bl*time+Centiloids_bl+meta_temporal_suvr_bl+mean_cortical_thickness_bl+hippo_volume_bl',long_cov,"(1|time|Subject_ID)",sep = "+")

# perform linear mixed effect regression
lmer(formula = mod4.3.5.2,data = data,contrasts=list(Gender = "contr.sum",Educ_Group=c(-1,1)))
```

- Examine whether sensory-motor system segregation at baseline predicts change in CDR-SOB score: 

```{r model 4.3.5.3}
# DV: CDR-SOB scores
# IV: sensory-motor system segregation at baseline
# Controlling for cortical Aβ (i.e., centiloids), meta-temporal tau deposition, mean grey matter cortical thickness and mean hippocampal volume at baseline, together with the longitudinal covariates

mod4.3.5.3 = paste('CDR_SOB ~ sensory_seg_bl*time+Centiloids_bl+meta_temporal_suvr_bl+mean_cortical_thickness_bl+hippo_volume_bl',long_cov,"(1|time|Subject_ID)",sep = "+")

# perform linear mixed effect regression
lmer(formula = mod4.3.5.3,data = data,contrasts=list(Gender = "contr.sum",Educ_Group=c(-1,1)))
```

- Examine whether senosry-motor system segregation at baseline predicts change in CDR score: 

```{r model 4.3.5.4}
# DV: CDR scores
# IV: sensory-motor system segregation at baseline
# Controlling for cortical Aβ (i.e., centiloids), meta-temporal tau deposition, mean grey matter cortical thickness and mean hippocampal volume at baseline, together with the longitudinal covariates

mod4.3.5.4 = paste('CDR ~ sensory_seg_bl*time+Centiloids_bl+meta_temporal_suvr_bl+mean_cortical_thickness_bl+hippo_volume_bl',long_cov,"(1|time|Subject_ID)",sep = "+")

# perform linear mixed effect regression
lmer(formula = mod4.3.5.4,data = data,contrasts=list(Gender = "contr.sum",Educ_Group=c(-1,1)))
```

2) Test the association between change in brain system segregation and change in cogntive status over time:

- Examine whether there is an interaction between change in brain system segregation and time on CDR-SOB scores: 

```{r model 4.3.5.5}
# DV: CDR-SOB scores
# IV: the interaction between changes in whole-brain system segregation and time
# controlling for the interaction between change in cortical Aβ (i.e., centiloids) and time, change in meta-temporal tau and time, change in mean grey matter cortical thickness and time, change in mean hippocampal volume and time, and the longitudinal covariates

mod4.3.5.5 = paste('CDR_SOB ~ delta_wb_seg*time+delta_centiloids*time+delta_meta_temporal_suvr*time+delta_thickness*time+delta_hippo_volume*time',long_cov,'+(1|PTID)',sep = "+")
            
lmer(formula = mod4.3.5.5,data = data,contrasts=list(Gender = "contr.sum",Educ_Group=c(-1,1))) 
```

- Examine whether there is an interaction between change in brain system segregation, age at baseline and time on CDR-SOB scores: 

```{r model 4.3.5.6}
# DV: CDR-SOB scores
# IV: the interaction between changes in whole-brain system segregation, age and time
# Controlling for change in cortical Aβ (i.e., centiloids) and time, change in meta-temporal tau deposition and time, change in mean grey matter cortical thickness and time, change in mean hippocampal volume and time

mod4.3.5.6 = paste('CDR_SOB ~ delta_wb_seg*time*Age_bl+delta_centiloids*time+delta_meta_temporal_suvr*time+delta_thickness*time+delta_hippo_volume*time',long_cov,'+(1|PTID)',sep = "+")

lmer(formula = mod4.3.5.6,data = data,contrasts=list(Gender = "contr.sum",Educ_Group=c(-1,1)))  
```

```{r model 4.3.5.7}
# DV: CDR-SOB scores
# IV: the interaction between changes in whole-brain system segregation, age and time

mod4.3.5.7 = paste('CDR_SOB ~ delta_wb_seg*time*Age_bl',long_cov,'+(1|PTID)',sep = "+")

lmer(formula = mod4.3.5.7,data = data,contrasts=list(Gender = "contr.sum",Educ_Group=c(-1,1)))  
```

3) Test the association between change in brain system segregation and change in cogntive status over time in individuals who are cognitively healthy at baseline (CDR = 0): 

```{r model 4.3.5.8}
# DV: CDR-SOB scores
# IV: the interaction between changes in whole-brain system segregation, age and time
# Controlling for change in cortical Aβ (i.e., centiloids) and time, change in meta-temporal tau deposition and time, change in mean grey matter cortical thickness and time, change in mean hippocampal volume and time

mod4.3.5.8 = paste('CDR_SOB ~ delta_wb_seg*time*Age_bl+delta_centiloids*time+delta_meta_temporal_suvr*time+delta_thickness*time+delta_hippo_volume*time',long_cov,'+(1|PTID)',sep = "+")

lmer(formula = mod4.3.5.8,data = data,contrasts=list(Gender = "contr.sum",Educ_Group=c(-1,1)))  
```

```{r model 4.3.5.9}
# DV: CDR-SOB scores
# IV: the interaction between changes in whole-brain system segregation, age and time

mod4.3.5.9 = paste('CDR_SOB ~ delta_wb_seg*time*Age_bl',long_cov,'+(1|PTID)',sep = "+")

lmer(formula = mod4.3.5.9,data = data,contrasts=list(Gender = "contr.sum",Educ_Group=c(-1,1)))  
```


