---
title: "Statistical Analysis Models (Main Text)"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
rm(list = ls())
```

# Resting-state system segregation underlies reserve in Alzheimer’s disease 


```{r define covariates}
# define covariates for cross-sectional analyses:
# age, self-reported gender, head motion (post-scrubbing FD)
cross_cov = 'Age+Gender+FD_postscrub'

# define covariates for longitudinal analyses: 
# age at baseline, self-reported gender, head motion (post-scrubbing FD) at baseline
long_cov = 'Age_bl+Gender+FD_postscrub_bl'
```

For analyses comparing cross-sectional differences in cognitive status/function and system segregation, all analyses included the cross-sectional analyses covariates indicated above.

For analyses comparing baseline system segregation and longitudinal changes in cognitive status/function, 
all analyses included the longitudinal analyses covariates indicated above.

## Functional brain network organization supports cognitive health among individuals with AD-related pathological burden

### Resting-state system segregation explains discrepancies in cognitive status among individuals with high amyloid burden

1) Examine difference in brain system segregation between cognitively unimpaired and cognitively impaired individuals designated as Aβ+: 

```{r model 1.1}
# DV: whole-brain system segregation
# IV: CDR group (CDR = 0 vs. CDR > 0)

mod1.1 <- paste('wb_seg ~ CDR_Group',cross_cov,sep = "+") 

# Perform ANCOVA
car::Anova(lm(formula = mod1.1,data = data,contrasts=
              list(CDR_Group="contr.treatment",Gender = "contr.sum")),type=3)
```

- Examine difference in brain system segregation between cognitively unimpaired and cognitively impaired individuals designated as Aβ+, while controlling for Centiloids: 

```{r model 1.2}
# DV: whole-brain system segregation
# IV: CDR group (CDR = 0 vs. CDR > 0)
# controlling for Centiloids and the primary covariates

mod1.2 <- paste('wb_seg ~ CDR_Group+Centiloids',cross_cov,sep = "+") 

# Perform ANCOVA
car::Anova(lm(formula = mod1.2,data = data,contrasts=
              list(CDR_Group="contr.treatment",Gender = "contr.sum")),type=3)

```

- Examine difference in system segregation between cognitively unimpaired and cognitively impaired individuals designated as Aβ+, while treating CDR as a continuous variable:

```{r model 1.3}
# DV: system segregation
# IV: CDR 
# controlling for the primary covariates and Centiloids

mod1.3 <- paste('wb_seg ~ CDR+Centiloids',cross_cov,sep = "+") 

# Perform multiple regression:
lm(formula = mod1.3,data = data,contrasts = list(Gender = "contr.sum")) 
```

2) Examine difference in system segregation between cognitively healthy and cognitively impaired individuals designated as Aβ+ and tau+: 

```{r model 1.4}
# DV: whole-brain system segregation
# IV: CDR group (CDR = 0 vs. CDR > 0)

mod1.4 <- paste('wb_seg ~ CDR_Group',cross_cov,sep = "+") 

# Perform ANCOVA
car::Anova(lm(formula = mod1.4,data = data,contrasts=
              list(CDR_Group="contr.treatment",Gender = "contr.sum")),type=3)

```

3) Examine whether system segregation moderates the relationship between tau and cognitive dysfunction

-Examine whether sensory-motor system segregation moderates the relationship between meta-temporal tau burden and CDR-SOB scores:

```{r model 1.5}
# DV: CDR-SOB
# IV: interaction between sensory-motor system segregation and meta-temporal tau uptake
mod1.5 = paste('CDR_SOB ~ sensory_seg*meta_temporal_suvr',cross_cov,sep = "+") 

#perform multiple linear regression
lm(formula = mod1.5,data = data,contrasts = list(Gender = "contr.sum")) 

```

- Examine whether association system segregation moderates the relationship between meta-temporal tau burden and CDR-SOB scores:

```{r model 1.6}
# DV: CDR-SOB
# IV: interaction between association system segregation and meta-temporal tau uptake
mod1.6 = paste('CDR_SOB ~ asso_seg*meta_temporal_suvr',cross_cov,sep = "+") 

#perform multiple linear regression
lm(formula = mod1.6,data = data,contrasts = list(Gender = "contr.sum")) 

```

### Resting-state system segregation relates to dementia severity, beyond the variance explained by AD-related pathology and brain structure

- Examine the main effects of system segregation, cortical Aβ (i.e., Centiloids), meta-temporal tau deposition, mean total grey matter cortical thickness and mean hippocampal volume on CDR scores: 

```{r model 2.1}
# DV: CDR scores
# IV: whole-brain system segregation, cortical Aβ (i.e., centiloids), meta-temporal tau deposition, mean total grey matter cortical thickness and mean hippocampal volume 

mod2.1= paste('CDR ~ wb_seg+Centiloids+meta_temporal_suvr+mean_cortical_thickness+hippo_volume',cross_cov,sep = "+") 

#perform multiple linear regression
lm(formula = mod2.1,data = data,contrasts = list(Gender = "contr.sum")) 

```

### The relationship between cognitive status and system segregation varies in relation to educational attainment 

- Examine the moderating effects of education on the relationship between system segregation and CDR scores, controlling for the effects of cortical Aβ (i.e., Centiloids), meta-temporal tau deposition, mean total grey matter cortical thickness, and mean hippocampal volume 

```{r model 3.1}
# DV: CDR scores
# IV: interaction between whole-brain system segregation and education 
mod2.1= paste('CDR ~ wb_seg*Educ_Group+Centiloids+meta_temporal_suvr+mean_cortical_thickness+hippo_volume',cross_cov,sep = "+") 

#perform multiple linear regression
lm(formula = mod2.1,data = data,contrasts = list(Gender = "contr.sum",Educ_Group=c(-1,1)))
```

## An individual’s resting-state system segregation predicts their cognitive decline over time, irrespective of their AD-related pathology and brain structure

1) Examine whether system segregation at baseline predicts change in CDR score: 

```{r model 3.1}
# DV: CDR scores
# IV: whole brain system segregation at baseline
# Controlling for cortical Aβ (i.e., centiloids), meta-temporal tau deposition, mean total grey matter cortical thickness and mean hippocampal volume at baseline, together with the longitudinal covariates

mod3.1 = paste('CDR ~ wb_seg_bl*time+Centiloids_bl+meta_temporal_suvr_bl+mean_cortical_thickness_bl+hippo_volume_bl',long_cov,"(1+time|Subject_ID)",sep = "+")

# perform linear mixed effect regression
lmer(formula = mod3.1,data = data,contrasts=list(Gender = "contr.sum"))
```


